---
title: "Impact of climate change on terrestrial vertebrates dynamics"
author: "Mathieu Pélissié"
date: "02/12/2021"
bibliography: biblio.bib
output: 
  html_document:
    toc: true
    theme: united
csl: nature.csl
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
#load targets
LPI.long <- tar_read(data_long)
LPI.mod <- tar_read(data_mod)
map_trends <- tar_read(map_trends)
mod_biome <- tar_read(mod_biome)
p_mod_biome <- tar_read(p_mod_biome)
mod_class <- tar_read(mod_class)
p_mod_class <- tar_read(p_mod_class)
```

# Introduction

According to the latest WWF report, the average population size of terrestrial 
vertebrates has fallen by 68%  In less than 50 years [@wwf_living_2020]. However, 
recent studies have even suggested that the dynamics of biodiversity are not 
really declining on average. They postulate that species at risk are dragging 
down global indices [@leung_clustered_2020], but that there are as many "winners" 
as "losers" species in total [@dornelas_balance_2019]. These last results have
mainly been documented through the BioTIME database [@dornelas_biotime_2018].
In this report, we propose to re-evaluate this through the Living Planet 
Database. We also aim to investigate whether  population  dynamics are
influenced  by  climate change.


# Population changes

## Methods

The database considered contains `r length(unique(LPI.long$species))` species of terrestrial vertebrates and
`r length(unique(LPI.long$id))` population time series distributed worldwide. We only focused on time 
series beginning after 1980 in order to be consistent with the climate data 
we analyse in the following part.

Each  population  time  series  was  fitted  with  a  linear regression  model  to  quantify  the  trend  through:

\begin{eqnarray*} 
X(t) = \alpha_{0} + \alpha_{1}t + \epsilon(t)
\end{eqnarray*}

Where $X$  is the abundance estimate, $t$ the time in years, with process error 
represented by $\epsilon$ ~ $N$(0,$\sigma$²).

## Results

The following map illustrates the distribution of the time series. Each 
point is colored according to its trend and sized according to the magnitude of
its trend.

```{r, echo=FALSE}
map_trends
```

Among the `r length(unique(LPI.long$id))`, `r sum(LPI.mod$slope_p<0.05 & LPI.mod$slope>0)` are
increasing, `r sum(LPI.mod$slope_p<0.05 & LPI.mod$slope<0)` are decreasing and 
`r sum(LPI.mod$slope_p>=0.05)` are showing constant trends (p>0.05).


# Population change drivers

## Biogeographical drivers (biome and class impacts).

We found biome signal among terrestrial vertebrates (p-value=`r p_mod_biome``),
the results are presented below.

```{r, echo=FALSE}
summary(mod_biome)
```

Surprisingly, no impact of the class belonging has been found (p-value=`r p_mod_class`).

## Climate change impacts


# References
